### 2021-03-07 math/ia_2021/paper.tex
final rp draft

$SS_{res} = \sum_{i} e_i^2 = \sum_{i} (y_i - p_i)^2$, \footnote{$\sum_{i=1}^n x_i$ means that all values of
    x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$
and its corresponding predicted value $p_i$.

@@

$SS_{tot} = \sum_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$,
or $\bar y = \frac{1}{n} \sum_{i=1}^n y_i$

>>>>

$SS_{res} = \sum\limits_{i} e_i^2 = \sum\limits_{i} (y_i - p_i)^2$, \footnote{$\sum\limits_{i=1}^n x_i$ means that all values of
    x are summed, beginning from $x_1$ and ending at $x_n$} where $e_i$ is the error term, equal to the difference of a data point $y_i$
and its corresponding predicted value $p_i$.

@@

$SS_{tot} = \sum\limits_{i} (y_i - \bar y)^2$, where $\bar y$ is the mean of the data $y_1,\dots, y_i$,
or $\bar y = \frac{1}{n} \sum\limits_{i=1}^n y_i$

### 2021-03-07 research project/sections/Analysis_of_results.tex
rp work
final rp draft

\pgfplotstableread{
    X Y
    5 0.7601666451
    10 1.288158894
    15 2.050116062 
    20 2.63982296 
    25 3.091802597
    30 3.819538355 
    35 4.442822933
    40 5.081979036 
    45 5.706284046 
    50 6.370962858 
    55 7.020516396 
    60 7.723451853   
}\vietetable

\pgfplotstableread{
    X Y
    5 0.5919837952 
    10 1.307063103  
    15 1.902475357  
    20 2.679748535  
    25 3.212277889  
    30 3.904948235  
    35 4.547913074  
    40 5.255579948  
    45 5.837199688  
    50 6.578524113  
    55 7.270073891  
    60 8.004565239

@@

\pgfplotstableread{
    X Y
    5 10
    10 17
    15 27
    20 35
    25 42
    30 51
    35 59
    40 67
    45 76
    50 84
    55 92
    60 101
}\vietetableiter

\pgfplotstableread{
    X Y
    5 9
    10 20
    15 29
    20 41
    25 50
    30 61
    35 71
    40 82
    45 91
    50 102
    55 112
    60 123
}\madhavatableiter

@@

The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
in the table shown below were found (rounded to 3 significant figures). The values for the decimal place 
0 have been dropped as it would not be logical to include them.

@@

\begin{table}[h]
    \noindent% <-- important
    \setlength\tabcolsep{3pt} % default: 6pt
    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
    \toprule
    Decimal place & Iterations needed, Viète's method (no unit) & Iterations needed, Madhava's method (no unit)\\
    \midrule
    5 & 10 & 9 \\
    10 & 17 & 20 \\
    15 & 27 & 29 \\
    20 & 35 & 41 \\
    25 & 42 & 50 \\
    30 & 51 & 61 \\
    35 & 59 & 71 \\
    40 & 67 & 82 \\
    45 & 76 & 91 \\
    50 & 84 & 102 \\
    55 & 92 & 112 \\
    60 & 101 & 123
    \end{tabularx}
\end{table}

@@

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
        title={Comparing the iterations needed for approximating the value of $\pi$},
        xlabel={Decimal places},
        ylabel={Iterations},
        xmin=0, xmax=60,
        ymin=0, ymax=130,
        ytick={0,10,20,30,40,50,60,70,80,90,100,110,120,130},
        xtick={0,10,20,30,40,50,60},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        scale=1.2
    ]

@@

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
        title={Comparing the time taken for approximating the value of $\pi$},
        xlabel={Decimal places},
        ylabel={Time taken [ms]},
        xmin=0, xmax=60,
        ymin=0, ymax=10,
        ytick={0,2,4,6,8,10},
        xtick={0,10,20,30,40,50,60},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        scale=1.2
    ]

@@

Where decimal places and iterations are compared, there is a clear difference between the two 
methods: Madhava's method indeed takes more iterations to converge to a specific decimal of 
the value $\pi$. However, in the second graph comparing time, it can be noted that Madhava's 
method is marginally faster until the decimal 20, when the two methods diverge. It could be 
assumed that this occurrence is due to differences between the types of the two methods. Madhava's 
method is what is called an alternating series, due to its nature of alternating between values in order 
to converge to a specific value, in this case $\pi$, while Viète's method is an infinite product. The 
difference between these two methods is underlined in (Figure 3) \ref{fig:comparaison}. It can be 
deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation 
accuracy increases in what looks like an exponential fashion.

\textit{More info to be added}

>>>>

\pgfplotstableread{
    X Y
    0.7601666451 5
    1.288158894 10
    2.050116062 15
    2.63982296 20
    3.091802597 25
    3.819538355 30
    4.442822933 35
    5.081979036 40
    5.706284046 45
    6.370962858 50
    7.020516396 55
    7.723451853 60
}\vietetable

\pgfplotstableread{
    X Y
    0.5919837952 5
    1.307063103 10
    1.902475357 15
    2.679748535 20
    3.212277889 25
    3.904948235 30
    4.547913074 35
    5.255579948 40
    5.837199688 45
    6.578524113 50
    7.270073891 55
    8.004565239 60

@@

\pgfplotstableread{
    X Y
    10  5
    17  10
    27  15
    35  20
    42  25
    51  30
    59  35
    67  40
    76  45
    84  50
    92  55
    101 60
}\vietetableiter

\pgfplotstableread{
    X Y
    9 5
    20 10
    29 15
    41 20
    50 25
    61 30
    71 35
    82 40
    91 45
    102 50
    112 55
    123 60
}\madhavatableiter

@@

The arithmetic mean of a decimal place was calculated for 100 trials and the data points
in the table shown below were found (rounded to 3 significant figures). The values for the decimal place
0 have been dropped as it would not be logical to include them.

@@

\begin{table}[h]
    \noindent% <-- important
    \setlength\tabcolsep{3pt} % default: 6pt
    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
        \toprule
        Decimal place & Iterations needed, Viète's method (no unit) & Iterations needed, Madhava's method (no unit) \\
        \midrule
        5             & 10                                          & 9                                             \\
        10            & 17                                          & 20                                            \\
        15            & 27                                          & 29                                            \\
        20            & 35                                          & 41                                            \\
        25            & 42                                          & 50                                            \\
        30            & 51                                          & 61                                            \\
        35            & 59                                          & 71                                            \\
        40            & 67                                          & 82                                            \\
        45            & 76                                          & 91                                            \\
        50            & 84                                          & 102                                           \\
        55            & 92                                          & 112                                           \\
        60            & 101                                         & 123
    \end{tabularx}
\end{table}

@@

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
            title={Comparing the iterations needed for approximating the value of $\pi$},
            xlabel={Iterations},
            ylabel={Decimal places},
            xmin=0, xmax=130,
            ymin=0, ymax=60,
            ytick={0,10,20,30,40,50,60},
            xtick={0,10,20,30,40,50,60,70,80,90,100,110,120,130},
            legend pos=south east,
            ymajorgrids=true,
            grid style=dashed,
            scale=1.2
        ]

@@

\begin{tikzpicture}[scale=1.1]
    \begin{axis}[
            title={Comparing the time taken for approximating the value of $\pi$},
            xlabel={Time taken [ms]},
            ylabel={Decimal places},
            xmin=0, xmax=10,
            ymin=0, ymax=60,
            ytick={0,10,20,30,40,50,60},
            xtick={0,2,4,6,8,10},
            legend pos=south east,
            ymajorgrids=true,
            grid style=dashed,
            scale=1.2
        ]

@@

Where decimal places and iterations are compared, there is a clear difference between the two
methods: Madhava's method indeed takes more iterations to converge to a specific decimal of
the value $\pi$. However, in the second graph comparing time, it can be noted that Madhava's
method is marginally faster until the decimal 20, when the lines of the two methods stray apart. It could be
assumed that this occurrence is due to differences between the types of the two methods. Although it 
could be argued that this difference is only due to random error, the results in this graph 
are those collected from the arithmetic mean over 100 trials, and so it can be assessed that this is not the case. Madhava's
method is what is called an alternating series, due to its nature of alternating between values in order
to converge to a specific value, in this case $\pi$, while Viète's method is an infinite product. The
difference between these two methods is underlined in (Figure 3) \ref{fig:comparaison}. It can be
deduced that at lower decimal values Madhava's method is faster as Viète's method's approximation
accuracy firstly increases in what seems similar to a curve from an exponential function $f(x) = k^x$.
It can also be said that while Madhava's method alternates between values such that $m_{n} > \pi, m_{n+1} < \pi$,
Viète's method $\pi_v$ approximates pi in a manner such that $\pi_v < \pi$.

### 2021-03-07 research project/sections/Background.tex
rp work
final rp draft

@@

  \begin{scriptsize}
  \draw [fill=xdxdff] (4.242640687119286,4.242640687119285) circle (2.5pt);
  \draw[color=xdxdff] (4.4772465116279085,4.623665116279071) node {$B$};
  \draw [fill=xdxdff] (-4.242640687119285,4.242640687119286) circle (2.5pt);
  \draw[color=xdxdff] (-4.295311627906972,4.623665116279071) node {$A$};
  \draw [fill=xdxdff] (-4.275936263984878,-4.209081736713964) circle (2.5pt);
  \draw[color=xdxdff] (-4.336241860465112,-4.55) node {$C$};
  \draw [fill=xdxdff] (6,0) circle (2.5pt);
  \draw[color=xdxdff] (6.237246511627907,0.3873860465116302) node {$E$};
  \draw [fill=xdxdff] (4.242640687119286,-4.242640687119285) circle (2.5pt);
  \draw[color=xdxdff] (4.3772465116279085,-4.5) node {$D$};
  \draw[color=black] (5.3595720930232575,2.3111069767441874) node {$H'$};
  \draw [fill=xdxdff] (5.12,2.12) circle (2.5pt);
  \draw[color=black] (3.947479069767444,-0.615404651162788) node {$H$};
  \draw [fill=xdxdff] (4.242640687119286,0) circle (2.5pt);
  \draw [fill=xdxdff] (0,0) circle (2.5pt);
  \draw[color=uuuuuu] (0.00003720930232879,-0.3464558139534907) node {$O$};
  \draw[color=qqwuqq] (1.5325953488372117,0.3873860465116302) node {$\alpha$};
  \end{scriptsize}
  \end{tikzpicture}
  \label{fig:image1}
  \caption[]{Circle with 1 segment from a $n$-gon with point $H$ and 2 
  segments from an $2n$-gon, one of which on point $H'$, inscribed in 
  a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
  \end{figure}

% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius of the circle with center $R = OB$,

@@

Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$, 
where $p$ is the perimeter of the polygon and $a$ is the apothem, in this case $BD \cdot n$ 
and $OH$ respectively, let $A_{n}$ equal the area of 
the polygon with $n$ sides such that: \\
$A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
$A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$

@@

\footnotetext{where $\prod$ signifies a product. Similar expression to $\sum$ }

\begin{figure}[h]

        \draw[xshift=-1cm] (1,-0.7) node[circle,fill,inner sep=1pt](a){} -- (5,-0.7) 
node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1} \boxed{\text{ iter. 1}}$](b){};

>>>>

\begin{figure}[H]

    \draw[xshift=-1cm] (1,-0.7) node[circle,fill,inner sep=1pt,label=left:$\boxed{\text{iter. 0}}$](a){} -- (5,-0.7)
    node[circle,fill,inner sep=1pt,label=right:$+\frac{4}{1} \boxed{\text{ iter. 1}}$](b){};

@@

    \begin{scriptsize}
      \draw [fill=xdxdff] (4.242640687119286,4.242640687119285) circle (2.5pt);
      \draw[color=xdxdff] (4.4772465116279085,4.623665116279071) node {$B$};
      \draw [fill=xdxdff] (-4.242640687119285,4.242640687119286) circle (2.5pt);
      \draw[color=xdxdff] (-4.295311627906972,4.623665116279071) node {$A$};
      \draw [fill=xdxdff] (-4.275936263984878,-4.209081736713964) circle (2.5pt);
      \draw[color=xdxdff] (-4.336241860465112,-4.55) node {$C$};
      \draw [fill=xdxdff] (6,0) circle (2.5pt);
      \draw[color=xdxdff] (6.237246511627907,0.3873860465116302) node {$E$};
      \draw [fill=xdxdff] (4.242640687119286,-4.242640687119285) circle (2.5pt);
      \draw[color=xdxdff] (4.3772465116279085,-4.5) node {$D$};
      \draw[color=black] (5.3595720930232575,2.3111069767441874) node {$H'$};
      \draw [fill=xdxdff] (5.12,2.12) circle (2.5pt);
      \draw[color=black] (3.947479069767444,-0.615404651162788) node {$H$};
      \draw [fill=xdxdff] (4.242640687119286,0) circle (2.5pt);
      \draw [fill=xdxdff] (0,0) circle (2.5pt);
      \draw[color=uuuuuu] (0.00003720930232879,-0.3464558139534907) node {$O$};
      \draw[color=qqwuqq] (1.5325953488372117,0.3873860465116302) node {$\alpha$};
    \end{scriptsize}
  \end{tikzpicture}
  \label{fig:image1}
  \caption[]{Circle with 1 segment from a $n$-gon (shown by the \textcolor{green}{green} dotted line) with point $H$ and 2
    segments from an $2n$-gon (shown by the \textcolor{red}{red} dotted line), one of which on point $H'$, inscribed in
    a circle of radius $OB$, adapted from (Boris Gourévitch) \cite{borisgourevitch2013}}
\end{figure}

% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius $R$ of the circle with center $O$ defined as $R = OB$,

@@

Since the equation for the area of a polygon is defined as $A = \frac{p \cdot a}{2}$,
where $p$ is the perimeter of the polygon and $a$ is the apothem\footnote{apothem: the segment 
from the center of a polygon to the midpoint of one of its sides, here $OH$ for the $n$-gon and 
$OH'$ for the $2n$-gon}, in this case $BD \cdot n$
and $OH$ respectively, let $A_{n}$ equal the area of
the polygon with $n$ sides (displayed in \textcolor{green}{green} in Figure 2) such that: \\
$A_{n} = \frac{OH \cdot BD \cdot n}{2}$ \\
$A_{n} = \frac{R \cos{\alpha} \cdot 2 R \sin{\alpha} \cdot n}{2} = n R^2 \sin{\alpha} \cos{\alpha}$

@@

\footnotetext{where $\prod x$ signifies a product. Similar expression to $\sum x$ as in the product of 
the terms $x$ are calculated, as opposed to the sum}

### 2021-03-07 research project/sections/Computational_approach.tex
rp work
final rp draft

The program assigns the time before the execution of the method to a variable \verb|t1| 
with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal 
places of the resultant approximation are counted and once a specified threshold is reached, 
a new \verb|t2| time variable is assigned and the time taken, defined by the difference between 
\verb|t2| and \verb|t1|, is stored. This process is repeated for all specified decimal accuracies and 
for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.

The \verb|mpmath| library was used for the floating point operations required 
for comparison between approximated values and the constant $\pi$ that wouldn't 
have been possible using standard Python libraries. \cite{mpmath}

>>>>

The program assigns the time before the execution of the method to a variable \verb|t1|
with the \verb|time.time()| Python function. At each iteration of the method, the number of valid decimal
places of the resultant approximation are counted. Once a specified decimal place is reached,
determined by a programming \verb|if| statement, that can be mathematically expressed as:

$| \pi - \pi_{approx} | < 10^{-n}$, where $\pi_{approx}$ is the approximation by the method and 
$n$ defines the number of correct decimal places desired.

A new \verb|t2| time variable is assigned and the time taken, defined by $\verb|t2| - \verb|t1|$, is stored. This process is repeated for all specified decimal accuracies and
for both methods. The times recorded were stored in a \verb|.csv| file for further analysis.

The \verb|mpmath| library was used for the floating point operations required
for comparison between approximated values and the constant $\pi$ that wouldn't
have been possible using standard Python libraries, due to floating-point limitations (computing 
term for decimal numbers). \cite{mpmath}

### 2021-03-07 research project/sections/Conclusion.tex
rp
rp work
final rp draft

To conclude, even though Viète's method can seem more convoluted in practice, due to its geometric origins, 
it can be seen through the experiment conducted in this paper that this is clearly not the case 
in a computational context. His method is simplified to a point where its geometric origins are no 
longer findable from its computational implementation. And despite the assumption that could be 
made based on the types of operations needed for each method, Viète's requiring arguably more 
computationally challenging operations. These operations, such as square-root function or the product function 
used on high-precision floating point numbers, can be said to be more demanding than what could be said 
is simple summation, in the case of Madhava. This essay could be considered in defense of 
what is known as geometric algebra --- which can clearly produce results more capable than ones 
originating from an analytical method.

@@

It must be noted that the experimental procedure in this paper is far from perfect. Firstly, 
there are many variables that would be difficult to control.

\textit{More info to be added}

>>>>

To conclude, even though Viète's method can seem more convoluted in practice, due to its geometric origins,
it can be seen through the experiment conducted in this paper that this is clearly not the case
in a computational context. His method is simplified to a point where its geometric origins are no
longer findable from its computational implementation. And despite the assumption that could be
made based on the types of operations needed for each method, Viète's requiring arguably more
computationally challenging operations. These operations, such as square-root function or the product function
used on high-precision floating point numbers, can be said to be more demanding than what could be said
is simple summation, in the case of Madhava.

An example of this kind of difference in computing power 
needed for different mathematical operations can be seen in the video-game \textit{Quake III Arena}, published in 1999. 
The inverse square root of a number needed to be calculated, however due to performance issues, the standard 
function could not be used. And as such, a new function for the approximation of the inverse square root of a number 
was programmed \cite{id-software_2012}, highly more efficient than the old one, which shows that a mathematical function 
implemented in computing can be more efficient than another.

Ultimately, a method 
of approximating the constant $\pi$ computationally can be more efficient than another, and to the 
somewhat surprising extent that this method is a geometric one --- many would think an analytical one, 
that uses simple summation as opposed to square roots, would be more efficient.

@@

The method used in this paper is not perfect, and at closer inspection and in 
retrospect, many issues can be found. Firstly, in the computational implementation of 
Madhava's method, the value $\sqrt{12}$ is calculated at every iteration, in order to 
determine the number of correct decimal places, which can 
lead to efficiency losses. Instead of this, the value could have been calculated once 
and stored in a variable, to be referenced at every iteration. However, after running 
the program again with the modified code, the results stayed similar.

Another weakness in the method used in this paper is that a higher number of trials 
could have been used, as well as the value of $\pi$ approximated to a higher 
number of decimal places. A higher number of decimal places was disregarded during the 
implementation of the methods as it was noticed that values over 60 would take an indeterminately 
long time. This should not have been overlooked, as a different kind of relationship 
could have been made apparent that way.

### 2021-03-07 research project/sections/Further_research.tex
research project
rp
rp work
final rp draft

\subsection{}

>>>>

\subsection{Investigating newer, more efficient algorithms}

As the computation of $\pi$ moved from paper to machines, as all did from accounting to
mathematics, newer and more efficient methods were developed, notably by Ramanujan and
the Chudnovsky brothers. These methods should also be investigated, as they 
are arguably more pertinent in this field.

\subsection{Investigating the uses of efficient code}

This paper does not discuss the many advantages of efficient code in detail, 
and a research on the speed of the mathematical functions in programming could 
be conducted. Computer program efficiency is always a pertinent topic because 
of a faster and faster world. An example of this could be the company Amazon, 
whose production speed has skyrocketed during the last decade. An increasingly 
fast demand also requires an increasingly fast computing infrastructure; to handle 
shipments, manage customers or even conduct job interviews (an experimental system that received 
major backlash in 2018). \cite{murad_2021}
